---
title: "Supplemental Material: Hierarchical structuring of weapon allometry in the European earwig."
authors:
  - name: Céline Ozkan
    affiliations:
      - ref: co
    corresponding: false
    email: celine.ozkan@etu.univ-lyon1.fr
    orcid: 
  - name: Clint D. Kelly
    affiliations:
      - ref: cdk
    corresponding: true
    email: kelly.clint@uqam.ca
    orcid: 0000-0002-0693-7211
    
affiliations:
  - id: cdk
    name: Département des Sciences biologiques, Université du Québec à Montréal, Montréal, Canada
  - id: co
    name: Université Claude Bernard Lyon 1, Lyon, France

filters:
  - authors-block

crossref:
  tbl-title: Table S
  fig-title: Figure S

format:
  docx:
    toc: false
    reference-doc: supp-template.docx
---

```{r echo=FALSE}
knitr::read_chunk('scripts/2026_earwig_allometry.R')
```

```{r analysis, include=FALSE,echo=FALSE, warning=FALSE, results="hide"}
```

\newpage


**Table S1.** Posterior summaries from the Bayesian sex-specific allometry model. The model estimated Gaussian regression of log forceps length including body size (logPc), sex, diet, density, and their interactions, with varying intercepts and slopes for mother identity (id_mere). Reported values are posterior medians and 95% credible intervals (CrI).
```{r}
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(stringr)
library(flextable)

post <- posterior_summary(mod_sex) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Parameter") %>%
  filter(!str_detect(Parameter, "lp__"))

post <- post %>%
  mutate(
    Section = case_when(
      str_detect(Parameter, "^b_") ~ "Fixed effects",
      str_detect(Parameter, "^sd_") ~ "Among-mother SD",
      str_detect(Parameter, "^cor_") ~ "Among-mother correlation (r)",
      str_detect(Parameter, "^sigma") ~ "Residual SD"
    )
  ) %>%
  filter(!is.na(Section))

clean_labels_sex <- function(param) {

  param <- str_replace(param, "^b_", "")
  param <- str_replace_all(param, "logPc", "Body size")
  param <- str_replace_all(param, "sexmale", "Sex (male)")
  param <- str_replace_all(param, "dietPOOR", "Diet (poor)")
  param <- str_replace_all(param, "density4", "Density = 4")
  param <- str_replace_all(param, "density8", "Density = 8")
  param <- str_replace_all(param, ":", " × ")
  param <- str_replace_all(param, "Intercept", "Intercept")

  param <- str_replace(param,
                       "sd_id_mere__Intercept",
                       "SD (Intercept)")
  param <- str_replace(param,
                       "sd_id_mere__logPc",
                       "SD (Body size slope)")
  param <- str_replace(param,
                       "cor_id_mere__Intercept__logPc",
                       "r (Intercept, Body size slope)")
  param <- str_replace(param,
                       "^sigma$",
                       "Residual SD")

  param
}

post$Parameter <- clean_labels_sex(post$Parameter)

final_table <- post %>%
  transmute(
    Section,
    Parameter,
    Estimate = round(Estimate, 2),
    `95% CrI` = paste0(round(Q2.5, 2), ", ", round(Q97.5, 2))
  ) %>%
  arrange(factor(Section,
                 levels = c("Fixed effects",
                            "Among-mother SD",
                            "Among-mother correlation (r)",
                            "Residual SD")))

flextable(final_table) %>%
  fontsize(size = 10, part = "all") %>%
  autofit()
```


**Table S2.** Posterior summaries from the Bayesian morph-specific male allometry model. The model estimated Gaussian regression of log forceps length including body size (logPc), morph group, diet, density, and their interactions, with varying intercepts and slopes for mother identity (id_mere). Reported values are posterior medians and 95% credible intervals (CrI).
```{r}
#| echo: false
#| message: false
#| warning: false

post <- posterior_summary(mod_morph) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Parameter") %>%
  filter(!str_detect(Parameter, "lp__"))

post <- post %>%
  mutate(
    Section = case_when(
      str_detect(Parameter, "^b_") ~ "Fixed effects",
      str_detect(Parameter, "^sd_") ~ "Among-mother SD",
      str_detect(Parameter, "^cor_") ~ "Among-mother correlation (r)",
      str_detect(Parameter, "^sigma") ~ "Residual SD"
    )
  ) %>%
  filter(!is.na(Section))

clean_labels_morph <- function(param) {

  param <- str_replace(param, "^b_", "")
  param <- str_replace_all(param, "logPc", "Body size")
  param <- str_replace_all(param, "morphmajor", "Morph (macrolabic)")
  param <- str_replace_all(param, "dietPOOR", "Diet (poor)")
  param <- str_replace_all(param, "density4", "Density = 4")
  param <- str_replace_all(param, "density8", "Density = 8")
  param <- str_replace_all(param, ":", " × ")
  param <- str_replace_all(param, "Intercept", "Intercept")

  param <- str_replace(param,
                       "sd_id_mere__Intercept",
                       "SD (Intercept)")
  param <- str_replace(param,
                       "sd_id_mere__logPc",
                       "SD (Body size slope)")
  param <- str_replace(param,
                       "cor_id_mere__Intercept__logPc",
                       "r (Intercept, Body size slope)")
  param <- str_replace(param,
                       "^sigma$",
                       "Residual SD")

  param
}

post$Parameter <- clean_labels_morph(post$Parameter)

final_table <- post %>%
  transmute(
    Section,
    Parameter,
    Estimate = round(Estimate, 2),
    `95% CrI` = paste0(round(Q2.5, 2), ", ", round(Q97.5, 2))
  ) %>%
  arrange(factor(Section,
                 levels = c("Fixed effects",
                            "Among-mother SD",
                            "Among-mother correlation (r)",
                            "Residual SD")))

flextable(final_table) %>%
  fontsize(size = 10, part = "all") %>%
  autofit()
```
