---
title: "Supplemental Material: Hierarchical structuring of weapon allometry in the European earwig."
authors:
  - name: Céline Ozkan
    affiliations:
      - ref: co
    corresponding: false
    email: celine.ozkan@etu.univ-lyon1.fr
    orcid: 
  - name: Clint D. Kelly
    affiliations:
      - ref: cdk
    corresponding: true
    email: kelly.clint@uqam.ca
    orcid: 0000-0002-0693-7211
    
affiliations:
  - id: cdk
    name: Département des Sciences biologiques, Université du Québec à Montréal, Montréal, Canada
  - id: co
    name: Université Claude Bernard Lyon 1, Lyon, France

filters:
  - authors-block

crossref:
  tbl-title: Table S
  fig-title: Figure S

format:
  docx:
    toc: false
    reference-doc: supp-template.docx
---

```{r echo=FALSE}
knitr::read_chunk('scripts/2026_earwig_allometry.R')
```

```{r analysis, include=FALSE,echo=FALSE, warning=FALSE, results="hide"}
```

\newpage

We modelled log-transformed forceps length as a Gaussian response with identity link. Fixed effects included all lower-order terms implied by interactions between body size (log pronotum width), sex or male morph, diet, and density. Maternal identity was included as a varying intercept and slope (on log body size), allowing both mean trait size and allometric slope to vary among families.


## Statistical models

### Sex difference model

We modelled log-transformed forceps length as:

$$
\log F_{ij} \sim \mathcal{N}(\mu_{ij}, \sigma)
$$

$$
\mu_{ij} = 
\beta_0 
+ \left( \log P_{ij} \times \text{sex}_{ij} \times \text{diet}_{ij} \times \text{density}_{ij} \right)
+ b_{0j} + b_{1j} \log P_{ij}
$$

where the interaction term expands to include all lower-order main effects and interactions among \(\log P\), sex, diet, and density.

Random effects (maternal identity):

$$
\begin{pmatrix}
b_{0j} \\
b_{1j}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
0 \\
0
\end{pmatrix},
\mathbf{\Sigma}
\right)
$$

$$
\mathbf{\Sigma} =
\begin{pmatrix}
\sigma^2_{0} & \rho\,\sigma_{0}\sigma_{1} \\
\rho\,\sigma_{0}\sigma_{1} & \sigma^2_{1}
\end{pmatrix}
$$

Priors:

$$
\beta_k \sim \mathcal{N}(0, 1), \quad
\beta_0 \sim \mathcal{N}(0, 2)
$$

$$
\sigma \sim t^+_3(0, 1), \quad
\sigma_0, \sigma_1 \sim t^+_3(0, 1), \quad
\rho \sim \text{LKJ}(2)
$$

---

### Male morph model

We modelled log-transformed forceps length as:

$$
\log F_{ij} \sim \mathcal{N}(\mu_{ij}, \sigma)
$$

$$
\mu_{ij} = 
\beta_0 
+ \left( \log P_{ij} \times \text{morph}_{ij} \times \text{diet}_{ij} \times \text{density}_{ij} \right)
+ b_{0j} + b_{1j} \log P_{ij}
$$

where the interaction term expands to include all lower-order main effects and interactions among \(\log P\), morph, diet, and density.

Random effects:

$$
\begin{pmatrix}
b_{0j} \\
b_{1j}
\end{pmatrix}
\sim \mathcal{N}
\left(
\begin{pmatrix}
0 \\
0
\end{pmatrix},
\mathbf{\Sigma}
\right)
$$

Priors:

$$
\beta_k \sim \mathcal{N}(0, 0.5), \quad
\beta_0 \sim \mathcal{N}(1.2, 0.5)
$$

$$
\sigma \sim t^+_3(0, 0.5), \quad
\sigma_0, \sigma_1 \sim t^+_3(0, 0.3), \quad
\rho \sim \text{LKJ}(2)
$$
\newpage

**Table S1.** Posterior summaries from the Bayesian sex-specific allometry model. The model estimated Gaussian regression of log forceps length including body size (logPc), sex, diet, density, and their interactions, with varying intercepts and slopes for mother identity (id_mere). Reported values are posterior medians and 95% credible intervals (CrI).
```{r}
#| echo: false
#| message: false
#| warning: false

library(dplyr)
library(stringr)
library(flextable)

post <- posterior_summary(mod_sex) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Parameter") %>%
  filter(!str_detect(Parameter, "lp__"))

post <- post %>%
  mutate(
    Section = case_when(
      str_detect(Parameter, "^b_") ~ "Fixed effects",
      str_detect(Parameter, "^sd_") ~ "Among-mother SD",
      str_detect(Parameter, "^cor_") ~ "Among-mother correlation (r)",
      str_detect(Parameter, "^sigma") ~ "Residual SD"
    )
  ) %>%
  filter(!is.na(Section))

clean_labels_sex <- function(param) {

  param <- str_replace(param, "^b_", "")
  param <- str_replace_all(param, "logPc", "Body size")
  param <- str_replace_all(param, "sexmale", "Sex (male)")
  param <- str_replace_all(param, "dietPOOR", "Diet (poor)")
  param <- str_replace_all(param, "density4", "Density = 4")
  param <- str_replace_all(param, "density8", "Density = 8")
  param <- str_replace_all(param, ":", " × ")
  param <- str_replace_all(param, "Intercept", "Intercept")

  param <- str_replace(param,
                       "sd_id_mere__Intercept",
                       "SD (Intercept)")
  param <- str_replace(param,
                       "sd_id_mere__logPc",
                       "SD (Body size slope)")
  param <- str_replace(param,
                       "cor_id_mere__Intercept__logPc",
                       "r (Intercept, Body size slope)")
  param <- str_replace(param,
                       "^sigma$",
                       "Residual SD")

  param
}

post$Parameter <- clean_labels_sex(post$Parameter)

final_table <- post %>%
  transmute(
    Section,
    Parameter,
    Estimate = round(Estimate, 2),
    `95% CrI` = paste0(round(Q2.5, 2), ", ", round(Q97.5, 2))
  ) %>%
  arrange(factor(Section,
                 levels = c("Fixed effects",
                            "Among-mother SD",
                            "Among-mother correlation (r)",
                            "Residual SD")))

flextable(final_table) %>%
  fontsize(size = 10, part = "all") %>%
  autofit()
```


**Table S2.** Posterior summaries from the Bayesian morph-specific male allometry model. The model estimated Gaussian regression of log forceps length including body size (logPc), morph group, diet, density, and their interactions, with varying intercepts and slopes for mother identity (id_mere). Reported values are posterior medians and 95% credible intervals (CrI).
```{r}
#| echo: false
#| message: false
#| warning: false

post <- posterior_summary(mod_morph) %>%
  as.data.frame() %>%
  tibble::rownames_to_column("Parameter") %>%
  filter(!str_detect(Parameter, "lp__"))

post <- post %>%
  mutate(
    Section = case_when(
      str_detect(Parameter, "^b_") ~ "Fixed effects",
      str_detect(Parameter, "^sd_") ~ "Among-mother SD",
      str_detect(Parameter, "^cor_") ~ "Among-mother correlation (r)",
      str_detect(Parameter, "^sigma") ~ "Residual SD"
    )
  ) %>%
  filter(!is.na(Section))

clean_labels_morph <- function(param) {

  param <- str_replace(param, "^b_", "")
  param <- str_replace_all(param, "logPc", "Body size")
  param <- str_replace_all(param, "morphmajor", "Morph (macrolabic)")
  param <- str_replace_all(param, "dietPOOR", "Diet (poor)")
  param <- str_replace_all(param, "density4", "Density = 4")
  param <- str_replace_all(param, "density8", "Density = 8")
  param <- str_replace_all(param, ":", " × ")
  param <- str_replace_all(param, "Intercept", "Intercept")

  param <- str_replace(param,
                       "sd_id_mere__Intercept",
                       "SD (Intercept)")
  param <- str_replace(param,
                       "sd_id_mere__logPc",
                       "SD (Body size slope)")
  param <- str_replace(param,
                       "cor_id_mere__Intercept__logPc",
                       "r (Intercept, Body size slope)")
  param <- str_replace(param,
                       "^sigma$",
                       "Residual SD")

  param
}

post$Parameter <- clean_labels_morph(post$Parameter)

final_table <- post %>%
  transmute(
    Section,
    Parameter,
    Estimate = round(Estimate, 2),
    `95% CrI` = paste0(round(Q2.5, 2), ", ", round(Q97.5, 2))
  ) %>%
  arrange(factor(Section,
                 levels = c("Fixed effects",
                            "Among-mother SD",
                            "Among-mother correlation (r)",
                            "Residual SD")))

flextable(final_table) %>%
  fontsize(size = 10, part = "all") %>%
  autofit()
```
